\documentclass[a4paper,10pt]{article}

%opening
\title{Reply to the Reviewers Regarding Paper 11}
\author{}

\begin{document}

\maketitle

\noindent Dear Reviewers, \\

\noindent First of all, we would like to thank you for your constructive comments regarding our paper.
To address your comments and suggested changes, we have made the following main revisions to the submitted manuscript:
\medskip

\noindent Reviewer 1:
\begin{itemize}

\item We have changed the manuscript title to ``Verifying Team Formation Protocols \emph{with Probabilistic Model Checking}", to reflect the fact that the paper introduces a novel approach (and its implementation) to analyse algorithms in a competitive setting using stochastic two-player games, rather than just performing analysis of distributed algorithms using existing techniques implemented in PRISM. We have also rewritten the contribution list (page 3) to reflect this.

\item We added a paragraph to the introduction (page 1, paragraph 3) explaining why \emph{probabilistic} model checking is needed to analyse this algorithm, highlighting all areas where probabilities occur. This includes, in particular, the case of the \textsc{Online} algorithm, where agents have to form optimal teams for the probabilistic \emph{distribution} of tasks (in this paper we used uniform, but any other distribution could be considered with only minor changes in the model description).

\item Throughout the paper, we clarified the use of synthesis of optimal strategies for agents, in particular paragraphs 1 and 5 of section 4 and paragraph 1 of section 5.

\item The meaning of ``hostile'' has been clarified throughout the paper by relating it to two player zero-sum games (e.g. paragraph 5 in section 1, paragraph 1 in section 5). Cooperative agents try to maximise the given reward whereas hostile ones try to minimise it. The solution of a stochastic two-player game (STPG) provides the ``optimal'' strategy for hostile agents which minimises the expectation of rewards. This can be interpreted as ``programming'' of hostile agents.

\item To comment on why this problem is \emph{not} a straightforward task allocation problem with multiple \emph{processes}, we would like to draw the reviewer's attention to the analysis of the competitive stochastic game setting, where \emph{processes} become \emph{players} having opposite goals and not yet having any pre-programmed execution strategies.

\end{itemize}

\noindent Reviewers 2 and 3:
\begin{itemize}

\item The typesetting has been revised.
\item We added a note to line 9 in Algorithms 1 and 2 to reflect the condition in the \textsc{Online} version of the algorithm.
\item More focus has been given to the analysis using bar charts rather than tables in the DTMC and MDP analysis sections.


\end{itemize}

\bigskip
\bigskip

\noindent Yours faithfully,\\


\noindent Taolue Chen

\noindent Marta Kwiatkowska

\noindent David Parker

\noindent Aistis Simaitis

\end{document}
